---
title: "Sp_analysis"
author: "Erin D'Agnese"
date: "4/14/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(tidyverse)
library(ggplot2)
library(here)
theme_set(theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")))
library(RColorBrewer)
library(jcolors)
#palette <- c('#e6194b', '#3cb44b', '#ffe119', '#4363d8', '#f58231', '#911eb4', '#46f0f0', '#f032e6', '#bcf60c', '#fabebe', '#008080', '#e6beff', '#9a6324', '#fffac8', '#800000', '#aaffc3', '#808000', '#ffd8b1', '#000075', '#808080', '#ffffff', '#000000')
```


```{r import data}
Sp_test_amp <- read.csv(here("20220615_Plamp_sens_amp.csv"))
Sp_test_map <- read.csv(here("20220615_Plamp_sens_map.csv"))

Pl_spec_amp <- read.csv(here("20220617_Plamp_spec_amp.csv"))
Pl_spec_map <- read.csv(here("20220617_Plamp_spec_map.csv"))

Rl_spec_amp <- read.csv(here("20220617_Rivlamp_spec_amp.csv"))
Rl_spec_map <- read.csv(here("20220617_Rivlamp_spec_map.csv"))

Rl_sens_amp <- read.csv(here("20220630_Rivlamp_sens_amp.csv"))
Rl_sens_map <- read.csv(here("20220630_Rivlamp_sens_map.csv"))

B_test_amp <- read.csv(here("20220608_Btrt_sens_amp.csv"))
B_test_map <- read.csv(here("20220608_Btrt_sens_map.csv"))

Bt_spec_amp <- read.csv(here("20220517_Btrt_ITS1_2_spec_amp.csv"))
Bt_spec_map <- read.csv(here("20220517_Btrt_ITS1_2_spec_map.csv"))

gb_amp <- read.csv(here("20220707_gBlock_amp.csv"))
gb_map <- read.csv(here("20220707_gBlock_map.csv"))
gb_copies <- read.csv(here("20220707_gBlock_copies.csv"))

gb_redo_amp <- read.csv(here("20220714_Bt_gblock_redo_amp.csv"))
gb_redo_map <- read.csv(here("20220714_Bt_gblock_redo_map.csv"))
```



```{r create a time column based on Ct}
Sp_test_amp$time <- Sp_test_amp$Cycle / 2
B_test_amp$time <- B_test_amp$Cycle / 2
Bt_spec_amp$time <- Bt_spec_amp$Cycle / 2
Pl_spec_amp$time <- Pl_spec_amp$Cycle / 2
Rl_spec_amp$time <- Rl_spec_amp$Cycle / 2
Rl_sens_amp$time <- Rl_sens_amp$Cycle / 2
gb_amp$time <- gb_amp$Cycle / 2
gb_redo_amp$time <- gb_redo_amp$Cycle / 2
```

format mapping file
```{r setting up the mapping and amp data into one df}
Sp_test_map %>% mutate(sample = paste(Sample.Name, Target.Name)) %>%
  subset(select= -c(Sample.Name,Target.Name)) %>% 
  subset( sample != " ") -> Sp_test_map

Pl_spec_map %>% mutate(sample = Sample.Name)  %>%
  subset(select= -c(Sample.Name))-> Pl_spec_map

Rl_spec_map %>% mutate(sample = Sample.Name)  %>%
  subset(select= -c(Sample.Name))-> Rl_spec_map

Rl_sens_map %>% mutate(sample = paste(Sample.Name, Target.Name)) %>%
  #subset(select= -c(Sample.Name,Target.Name)) %>% 
  subset( sample != " ")-> Rl_sens_map

B_test_map %>% 
  mutate(sample = paste(Sample.Name, Target.Name)) %>%
  subset(select= -c(Sample.Name,Target.Name)) %>% 
  subset( sample != " ") -> B_test_map

Bt_spec_map %>% mutate(sample = Sample.Name)  %>%
  subset(select= -c(Sample.Name))-> Bt_spec_map

gb_map %>%
  mutate(sample = paste(Sample.Name, Target.Name)) %>%
  subset(select= -c(Sample.Name,Target.Name)) %>% 
  subset( sample != " ") -> gb_map

gb_redo_map %>%
  mutate(sample = paste(Sample.Name, Target.Name)) %>%
  subset(select= -c(Sample.Name,Target.Name)) %>% 
  subset( sample != " ") -> gb_redo_map

gb_copies %>%
  mutate(sample = paste(Sample.Name, Target.Name)) %>%
  subset( sample != " ") -> gb_copies


```


```{r fixing dataframes}
inner_join(gb_map, gb_copies, by = "sample") -> gb_map2
inner_join(gb_redo_map, gb_copies, by = "sample") -> gb_redo_map2
```

```{r}
inner_join(Sp_test_amp, Sp_test_map, by = "Well") -> Sp_test
inner_join(B_test_amp, B_test_map, by = "Well") -> Bt_test
inner_join(Bt_spec_amp, Bt_spec_map, by = "Well") -> Bt_spec
inner_join(Pl_spec_amp, Pl_spec_map, by = "Well") -> Pl_spec
inner_join(Rl_spec_amp, Rl_spec_map, by = "Well") -> Rl_spec
inner_join(Rl_sens_amp, Rl_sens_map, by = "Well") -> Rl_sens
inner_join(gb_amp, gb_map2, by = "Well") -> gb_sens
inner_join(gb_redo_amp, gb_redo_map2, by = "Well") -> gb_redo_sens

Sp_test %>% mutate(across(where(is.character), str_remove_all, pattern = fixed(" "))) -> Sp_test

Bt_test %>% mutate(across(where(is.character), str_remove_all, pattern = fixed(" "))) -> Bt_test

Bt_spec %>% mutate(across(where(is.character), str_remove_all, pattern = fixed(" "))) -> Bt_spec

Pl_spec %>% mutate(across(where(is.character), str_remove_all, pattern = fixed(" "))) %>% drop_na() -> Pl_spec

Rl_spec %>% mutate(across(where(is.character), 
                          str_remove_all, pattern = fixed(" "))) %>%
  drop_na() -> Rl_spec 

Rl_sens %>% mutate(across(where(is.character), str_remove_all, pattern = fixed(" "))) -> Rl_sens 

gb_sens %>% mutate(across(where(is.character), str_remove_all, pattern = fixed(" "))) -> gb_sens

gb_redo_sens %>% mutate(across(where(is.character), str_remove_all, pattern = fixed(" "))) -> gb_redo_sens
```

```{r}
ggplot(gb_sens, aes(x=Cycle, y=FAM, color = Sample.Name, shape= Target.Name)) + geom_point()
```

```{r}
ggplot(Bt_test, aes(x=Cycle, y=FAM, color = sample)) + geom_point()

```

Make a df for each species in the Ggblock trail
```{r}
gb_sens %>%
  filter(str_detect(sample, "Etri")) -> gb_Plamp
gb_sens %>%
  filter(str_detect(sample, "S.con")) -> gb_Btrt
gb_sens%>%
  filter(str_detect(sample, "L.ayr")) -> gb_Rlamp


```


calculating background from NTCs
```{r pull out NTC values to calc background}
#take average of NTC

Sp_test %>% filter(sample == "NTC") -> Sp_test_NTC
Bt_test %>% filter(sample == "NTC") -> Bt_test_NTC
Rl_spec %>% filter(sample == "NTC") -> Rl_spec_NTC
Rl_sens %>% filter(sample == "NTC") -> Rl_sens_NTC
gb_Btrt %>% filter(Sample.Name == "NTC") -> gb_Btrt_NTC
gb_redo_sens %>% filter(Sample.Name == "NTC") -> gb_redo_sens_NTC
```

```{r using the function to calc background for correction}
#made this a function 'background_fl' but requires a NTC subset
#as.tbl(Sp_sens1_NTC) %>% 
#  bind_rows(mutate(Sp_sens1_NTC, Cycle=NA, sample=NA)) %>% 
#  bind_rows(mutate(Sp_sens1_NTC, Well=NA)) %>% 
#  group_by(Cycle) %>% 
#  summarise_all(c("mean", "sd")) %>%
#  subset(select = c(Cycle, FAM_mean, FAM_sd)) -> Sp_sens1_NTC_summary

#Sp_sens1_NTC_summary %>% rename(BGRn_mean = Rn_mean) %>%
#  rename(BGRn_sd = Rn_sd) -> Sp_sens1_background

source(here("background_fl.R"))


Sp_test_NTC_bg <- background_fl(Sp_test_NTC)

Bt_test_NTC_bg <- background_fl(Bt_test_NTC)

Rl_spec_NTC_bg <- background_fl(Rl_spec_NTC)

Rl_sens_NTC_bg <- background_fl(Rl_sens_NTC)

gb_Btrt_NTC_bg <- background_fl(gb_Btrt_NTC)

gb_redo_sens_NTC_bg <- background_fl(gb_redo_sens_NTC)


#join the mean values with the cycle numbers in the main data

left_join(Sp_test, Sp_test_NTC_bg, by = "Cycle") -> Sp_test_total

left_join(Bt_test, Bt_test_NTC_bg, by = "Cycle") -> Bt_test_total

left_join(Rl_spec, Rl_spec_NTC_bg, by = "Cycle") -> Rl_spec_total

left_join(Rl_sens, Rl_sens_NTC_bg, by = "Cycle") -> Rl_sens_total

left_join(gb_Btrt, gb_Btrt_NTC_bg, by = "Cycle") -> gb_Btrt_total

```




Let's try finding the derivative for each sample and the NTC by using the cumulative sum of the 
change in Y over the change in x - use this to determine detections/non-detections for the specificity
```{r derivative}
Pl_spec %>%
  group_by(Well) %>%
  mutate(FAMcs = cumsum(FAM)) %>%
  mutate(FAMfinal = last(FAMcs))-> Pl_spec_FAMcs

Pl_spec_FAMcs %>%
  filter(!str_detect(sample, "GR")) -> Pl_spec_tmp

Pl_spec_FAMcs %>%   
  select(Well, sample, FAMcs) %>%
  group_by(Well, sample) %>%
  summarise(FAMfinal = sum(FAMcs)) %>%
  mutate(detect = case_when(FAMfinal > 0 ~"Yes",
                            FAMfinal <= 0 ~ "No")) -> Pl_spec_detect




ggplot(Pl_spec_tmp, aes(x=time, y=FAMcs, color=sample)) + geom_point() +
  ylab(expression("Cumulative Sum of change of Fluorescence"~(Delta *"RFU"))) +
  xlab("Time (min)")



```

```{r}
ggplot(Pl_spec_tmp, aes(x=time, y=FAM, color=sample)) +
  geom_smooth()+
  geom_hline(yintercept = thresh, linetype = "dashed", color = "grey36", size = 1) +
  annotate("text", x = 60, y= thresh, vjust = 1, label = "Threshold", size = 2.5) +
  ylab("Flourescence Intensity") +
  xlab("Time (min)")
```


```{r}
Rl_spec %>% group_by(Well) %>%
  mutate(FAMcs = cumsum(FAM)) %>%
  mutate(FAMfinal = last(FAMcs))-> Rl_spec_FAMcs

Rl_spec_FAMcs %>%   
  select(Well, sample, FAMcs) %>%
  group_by(Well, sample) %>%
  summarise(FAMfinal = sum(FAMcs)) %>%
  mutate(detect = case_when(FAMfinal > 0 ~"Yes",
                            FAMfinal <= 0 ~ "No")) -> Rl_spec_detect


Rl_spec_FAMcs %>%
  filter(!str_detect(sample, "GR")) -> Rl_spec_tmp

ggplot(Rl_spec_tmp, aes(x=time, y=FAMcs, color=sample)) + geom_point() +
  ylab(expression("Cumulative Sum of change of Fluorescence"~(Delta *"RFU"))) +
  xlab("Time (min)")



```

```{r}
ggplot(Rl_spec_tmp, aes(x=time, y=FAM, color=sample)) +
  geom_smooth()+
  geom_hline(yintercept = thresh, linetype = "dashed", color = "grey36", size = 1) +
  annotate("text", x = 60, y= thresh, vjust = 1, label = "Threshold", size = 2.5) +
  ylab("Flourescence Intensity") +
  xlab("Time (min)")
```


```{r}
Bt_spec %>%
  group_by(Well) %>%
  mutate(FAMcs = cumsum(FAM)) %>%
  mutate(FAMfinal = last(FAMcs))-> Bt_spec_FAMcs

Bt_spec_FAMcs %>%   
  select(Well, sample, FAMcs) %>%
  group_by(Well, sample) %>%
  summarise(FAMfinal = sum(FAMcs)) %>%
  mutate(detect = case_when(FAMfinal > 0 ~"Yes",
                            FAMfinal <= 0 ~ "No")) -> Bt_spec_detect

ggplot(Bt_spec_FAMcs, aes(x=time, y=FAMcs, color=sample)) + geom_point() +
  ylab(expression("Cumulative Sum of change of Fluorescence"~(Delta *"RFU"))) +
  xlab("Time (min)")

```
specificity plot with zero as the threshold
```{r}
ggplot(Bt_spec, aes(x=time, y=FAM, color=sample)) +
  geom_smooth()+
  geom_hline(yintercept = thresh, linetype = "dashed", color = "grey36", size = 1) +
  annotate("text", x = 60, y= thresh, vjust = 1, label = "Threshold", size = 2.5) +
  ylab("Flourescence Intensity") +
  xlab("Time (min)")
```



```{r}
#let's do it by assay
Sp_test_FAMcs %>% 
  filter(str_detect(Target.Name, "P.lamp_crRNA_1")) -> Plamp_FAMcs
Sp_test_FAMcs %>% 
  filter(str_detect(Target.Name, "R.lamp_crRNA_1")) -> Rlamp_FAMcs
#Sp_test_FAMcs %>% 
#  filter(str_detect(Target.Name, "B.trt_ITS1_crRNA_1")) -> Btrt_ITS1_1_FAMcs
#Sp_test_FAMcs %>% 
#  filter(str_detect(Target.Name, "B.trt_ITS1_crRNA_2")) -> Btrt_ITS1_2_FAMcs
#Sp_test_FAMcs %>% 
#  filter(str_detect(Target.Name, "B.trt_Wilcox_crRNA_1")) -> Btrt_Wil_FAMcs

Plamp_FAMcs %>%   
  select(Well, sample, FAMcs) %>%
  group_by(Well, sample) %>%
  summarise(FAMfinal = sum(FAMcs)) %>%
  mutate(detect = case_when(FAMfinal > 0 ~"Yes",
                            FAMfinal <= 0 ~ "No")) -> Plamp_test_detect

Rlamp_FAMcs %>%   
  select(Well, sample, FAMcs) %>%
  group_by(Well, sample) %>%
  summarise(FAMfinal = sum(FAMcs)) %>%
  mutate(detect = case_when(FAMfinal > 0 ~"Yes",
                            FAMfinal <= 0 ~ "No")) -> Rlamp_test_detect

```


```{r}
ggplot(Plamp_FAMcs, aes(x=Cycle, y=FAMcs, color=sample)) + geom_point() +
 ylab(expression("Cumulative Sum of change of Fluorescence"~(Delta *"RFU"))) +
  xlab("Time (min)")
    #+ geom_smooth()
```
So there is one R.lamprey species that is amplifying with this assay

```{r}
ggplot(Plamp_FAMcs, aes(x=Cycle, y=FAM, color=sample)) + geom_point() #+ geom_smooth()

```


```{r}
ggplot(Rlamp_FAMcs, aes(x=Cycle, y=FAMcs, color=sample)) + geom_point() +
 ylab(expression("Cumulative Sum of change of Fluorescence"~(Delta *"RFU"))) +
  xlab("Time (min)")

```

This one has 2 of the 3 river lamprey and nothing else amplifying! Yahoo

```{r}
ggplot(Rlamp_FAMcs, aes(x=Cycle, y=FAM, color=sample)) + geom_point() #+ geom_smooth()

```

```{r}
ggplot(Btrt_ITS1_1_FAMcs, aes(x=Cycle, y=FAMcs, color=sample)) + geom_point() #+ geom_smooth()

```
Two of the three Bull trout amplified well and the third one if the only other isolate that also worked!

```{r}
ggplot(Btrt_ITS1_1_FAMcs, aes(x=Cycle, y=FAM, color=sample)) + geom_point() #+ geom_smooth()

```

```{r}
ggplot(Btrt_ITS1_2_FAMcs, aes(x=Cycle, y=FAMcs, color=sample)) + geom_point() #+ geom_smooth()

```
This one only has Bull trout working as well!

```{r}
ggplot(Btrt_ITS1_2_FAMcs, aes(x=Cycle, y=FAM, color=sample)) + geom_point() #+ geom_smooth()
```



```{r}
ggplot(Btrt_Wil_FAMcs, aes(x=Cycle, y=FAMcs, color=sample)) + geom_point() #+ geom_smooth()
```
Oh darn, the Dolly varden also amplified


```{r}
ggplot(Btrt_Wil_FAMcs, aes(x=Cycle, y=FAM, color=sample)) + geom_point() #+ geom_smooth()
```


```{r}

```
One of the Bull trout amplified well, the other two also had positive slopes and nothing else did

```{r}

```


combine the detections and export to make a table with metadata
```{r}
rbind(Sp_eDNA1_detections, Sp_eDNA2_detections) -> Sp_eDNA_detections

inner_join(Sp_eDNA_detections,Sp_qPCR_results, keep=TRUE) -> Sp_eDNA_qPCR

write.csv(Sp_eDNA_detections, "Sp_Carson_eDNA_cumsum.csv", row.names = FALSE)

Sp_eDNA_qPCR %>%
  mutate(qPCR_det = case_when(CT == "No Amp" ~ "FALSE",
                              CT!= "No Amp" ~ "TRUE")) -> Sp_eDNA_qPCR
         

ggplot(Sp_eDNA_qPCR, aes(x=Quantity, y=FAMfinal))+
  geom_point()

```



Find the highest threshold for the sensitivity data
```{r}
Sp_test_total %>% 
  mutate(CorFl = FAM - BG_cor) %>%
  mutate(CycleTh = ((10*BG_sd) + BG_mean)) %>%
  mutate(CorCt = case_when(CorFl > CycleTh ~ "Y", 
         CorFl <= CycleTh ~ "N")) -> Sp_test_BgCor


MaxTh <- max(Sp_test_BgCor$CycleTh)  
AvgTh <- mean(Sp_test_BgCor$CycleTh)

#if the MaxTh is a negative value assign it to zero
MaxTh[MaxTh <= 0] <- 0

#use this as the threshold in the graph
#make a BgCor detection based on when it crosses MaxTh
Sp_test_BgCor %>% 
  mutate(MaxCycTh = case_when(CorFl > MaxTh ~ "Y",
                              CorFl <= MaxTh ~ "N")) -> Sp_test_BgCor
Sp_test_BgCor %>% 
  mutate



Bt_test_total %>% 
  mutate(CorFl = FAM - BG_cor) %>%
  mutate(CycleTh = ((10*BG_sd) + BG_mean)) %>%
  mutate(CorCt = case_when(CorFl > CycleTh ~ "Y", 
         CorFl <= CycleTh ~ "N")) -> Bt_BgCor

MaxTh <- max(Bt_BgCor$CycleTh)
AvgTh <- mean(Bt_BgCor$CycleTh)

gb_Btrt_total %>% 
  mutate(CorFl = FAM - BG_cor) %>%
  mutate(CycleTh = ((10*BG_sd) + BG_mean)) %>%
  mutate(CorCt = case_when(CorFl > CycleTh ~ "Y", 
         CorFl <= CycleTh ~ "N")) -> gb_Btrt_Bgcor

MaxTh <- max(gb_Btrt_Bgcor$CycleTh)
AvgTh <- mean(gb_Btrt_Bgcor$CycleTh)
```


Using 0 as the threshold
```{r}
Rl_spec_total %>%
   mutate(Detect = case_when(FAM > 0 ~ "Y", 
         FAM <= 0 ~ "N")) -> Rl_spec_Ct

Sp_test_total %>%
   mutate(Detect = case_when(FAM > 0 ~ "Y", 
         FAM <= 0 ~ "N")) -> Sp_test_Ct

Bt_test_total %>%
   mutate(Detect = case_when(FAM > 0 ~ "Y", 
         FAM <= 0 ~ "N")) -> Bt_test_Ct
thresh = 0  

Rl_sens_total %>% 
  mutate(Detect = case_when(FAM > 0 ~ "Y", 
         FAM <= 0 ~ "N")) -> Rl_sens_Ct

gb_Btrt %>% mutate(Detect = case_when(FAM > 0 ~ 1, 
         FAM <= 0 ~ 0)) -> gb_Btrt_Ct

gb_Plamp %>% mutate(Detect = case_when(FAM > 0 ~ 1, 
         FAM <= 0 ~ 0)) -> gb_Plamp_Ct

gb_Rlamp %>% mutate(Detect = case_when(FAM > 0 ~ 1, 
         FAM <= 0 ~ 0)) -> gb_Rlamp_Ct

gb_redo_sens %>% mutate(Detect = case_when(FAM > 0 ~ 1, 
         FAM <= 0 ~ 0)) -> gb_Btrt_redo_Ct

```

Bull trout sensitivity plot
```{r}
ggplot(Bt_test_Ct, aes(x=time, y=FAM, color=sample)) +
  geom_smooth()+
  geom_hline(yintercept = thresh, linetype = "dashed", color = "grey36", size = 1) +
  annotate("text", x = 60, y= thresh, vjust = 1, label = "Threshold", size = 2.5) +
  ylab("Flourescence Intensity") +
  xlab("Time (min)") +
  ggtitle("Bull trout tissue sensitivity test")
```

Pacific lamprey sensitivity plot
```{r}
ggplot(Sp_test_Ct, aes(x=time, y=FAM, color=sample)) +
  geom_smooth()+
  geom_hline(yintercept = thresh, linetype = "dashed", color = "grey36", size = 1) +
  annotate("text", x = 60, y= thresh, vjust = 1, label = "Threshold", size = 2.5) +
  ylab("Flourescence Intensity") +
  xlab("Time (min)")
```


River lamprey sensitivity plot
```{r}
ggplot(Rl_sens_Ct, aes(x=time, y=FAM, color=sample)) +
  geom_smooth()+
  geom_hline(yintercept = thresh, linetype = "dashed", color = "grey36", size = 1) +
  annotate("text", x = 60, y= thresh, vjust = 1, label = "Threshold", size = 2.5) +
  ylab("Flourescence Intensity") +
  xlab("Time (min)") 

```

Gblock sensitivity plots
```{r}
ggplot(gb_Btrt_Ct, aes(x=time, y=FAM, color=sample)) +
  geom_smooth()+
  geom_hline(yintercept = thresh, linetype = "dashed", color = "grey36", size = 1) +
  annotate("text", x = 60, y= thresh, vjust = 1, label = "Threshold", size = 2.5) +
  ylab("Flourescence Intensity") +
  xlab("Time (min)") +
  ggtitle("Initial Bull trout gblock test")

```

```{r}
ggplot(gb_Btrt_redo_Ct, aes(x=time, y=FAM, color=sample)) +
  geom_smooth()+
  geom_hline(yintercept = thresh, linetype = "dashed", color = "grey36", size = 1) +
  annotate("text", x = 60, y= thresh, vjust = 1, label = "Threshold", size = 2.5) +
  ylab("Flourescence Intensity") +
  xlab("Time (min)") +
  ggtitle("Redo of Bull trout gBlock test")
```
plot the background corrected FAM for the gblock Bull trout
```{r}


ggplot(gb_Btrt_Bgcor, aes(x=time, y=CorFl, color=sample)) +
  geom_smooth()+
  geom_hline(yintercept = AvgTh, linetype = "dashed", color = "grey36", size = 1) +
  annotate("text", x = 60, y= thresh, vjust = 1, label = "Threshold", size = 2.5) +
  ylab("Flourescence Intensity") +
  xlab("Time (min)") +
  ggtitle("Background corrected Initial Bull trout gblock test")
```


```{r}
ggplot(gb_Plamp_Ct, aes(x=time, y=FAM, color=sample)) +
  geom_smooth()+
  geom_hline(yintercept = thresh, linetype = "dashed", color = "grey36", size = 1) +
  annotate("text", x = 60, y= thresh, vjust = 1, label = "Threshold", size = 2.5) +
  ylab("Flourescence Intensity") +
  xlab("Time (min)")
```
river lamprey gblock plot
```{r}
ggplot(gb_Rlamp_Ct, aes(x=time, y=FAM, color=sample)) +
  geom_smooth()+
  geom_hline(yintercept = thresh, linetype = "dashed", color = "grey36", size = 1) +
  annotate("text", x = 60, y= thresh, vjust = 1, label = "Threshold", size = 2.5) +
  ylab("Flourescence Intensity") +
  xlab("Time (min)")
```

```{r}
#find the first time a sample crosses the 0 threshold
Rl_sens_total %>%
  mutate(tissue = case_when(
    str_detect(sample, "L.ayr_2") ~ "L.ayresii - UWBM",
    str_detect(sample, "NOAA") ~ "L.ayresii - NOAA",
    str_detect(sample, "Sk5-27") ~ "L.ayresii - Skagit-USGS 5-27"
  )) -> Rlamp_sens_total

Rlamp_sens_total %>% 
  group_by(sample) %>%
  #filter(FAM > 0) %>%
  mutate(CtPos = FAM > 0) %>%
  filter(CtPos == TRUE) %>%
  filter(time == min(time)) %>%
  slice(1) %>% # takes the first occurrence if there is a tie
  ungroup() -> Rlamp_sens_Ct

ggplot(Rlamp_sens_Ct, aes(x=time, y=log(conc), color=tissue)) + 
  geom_point() +
  geom_line() 


```

do that for each gBlock with copy number
```{r}
gb_Btrt_Bgcor %>% 
  group_by(sample) %>%
  #filter(FAM > 0) %>%
  mutate(CtPos = CorFl > AvgTh) %>%
  filter(CtPos == TRUE) %>%
  filter(time == min(time)) %>%
  slice(1) %>% # takes the first occurrence if there is a tie
  ungroup() -> gb_Btrt_1

ggplot(gb_Btrt_1, aes(x=time, y=log(copies.ul), color=Sample.Name)) + 
  geom_point(alpha = 4) 
  #geom_line() 

```

```{r}
gb_Plamp_Ct %>% 
  group_by(sample) %>%
  #filter(FAM > 0) %>%
  mutate(CtPos = FAM > 0) %>%
  filter(CtPos == TRUE) %>%
  filter(time == min(time)) %>%
  slice(1) %>% # takes the first occurrence if there is a tie
  ungroup() -> gb_Plamp_1

ggplot(gb_Plamp_1, aes(x=time, y=log(copies.ul), color=Sample.Name)) + 
  geom_point(alpha = 4) 
  #geom_line() 

```

```{r}
gb_Rlamp_Ct %>% 
  group_by(sample) %>%
  #filter(FAM > 0) %>%
  mutate(CtPos = FAM > 0) %>%
  filter(CtPos == TRUE) %>%
  filter(time == min(time)) %>%
  slice(1) %>% # takes the first occurrence if there is a tie
  ungroup() -> gb_Rlamp_1

ggplot(gb_Rlamp_1, aes(x=time, y=log(copies.ul), color=Sample.Name)) + 
  geom_point(alpha = 4) 
  #geom_line() 
```



```{r average across the TR in the sensitivity trial after BG correction}
Sp_test_BgCor %>%
  select(Well, Cycle, sample, CorFl) -> Sp_test_tmp

Sp_test_tmp %>%
  group_by(Cycle, sample) %>% 
  summarise(AvgCorFl = mean(CorFl, na.rm=TRUE), SdCorFl = sd(CorFl))-> Sp_test_tmp

left_join(Sp_test_BgCor, Sp_test_tmp, by=c("sample","Cycle")) -> Sp_test_all
```

```{r}
Plamp_BgCor %>%
  select(Well, Cycle, sample, CorFl) -> Plamp_tmp

Plamp_tmp %>%
  group_by(Cycle, sample) %>% 
  summarise(AvgCorFl = mean(CorFl, na.rm=TRUE), SdCorFl = sd(CorFl))-> Plamp_tmp

left_join(Plamp_BgCor, Plamp_tmp, by=c("sample","Cycle")) -> Plamp_all

```

```{r}
Rlamp_BgCor %>%
  select(Well, Cycle, sample, CorFl) -> Rlamp_tmp

Rlamp_tmp %>%
  group_by(Cycle, sample) %>% 
  summarise(AvgCorFl = mean(CorFl, na.rm=TRUE), SdCorFl = sd(CorFl))-> Rlamp_tmp

left_join(Rlamp_BgCor, Rlamp_tmp, by=c("sample","Cycle")) -> Rlamp_all
```

```{r}
Bt_BgCor %>%
  select(Well, Cycle, sample, CorFl) -> Bt_tmp

Bt_tmp %>%
  group_by(Cycle, sample) %>% 
  summarise(AvgCorFl = mean(CorFl, na.rm=TRUE), SdCorFl = sd(CorFl))-> Bt_tmp

left_join(Bt_BgCor, Bt_tmp, by=c("sample","Cycle")) -> Bt_all

```



let's just plot one and see what it looks like
```{r}
ggplot(Sp_test_tmp, aes(x=Cycle, y=AvgCorFl, color=sample)) +
  geom_point() 

```

```{r}
Bt_tmp %>%
  filter(str_detect(sample, "S.con-1")) -> Bt_tmp_1
ggplot(Bt_tmp, aes(x=Cycle, y=AvgCorFl, color=sample)) +
  geom_point() 
```


let's plot ones with the threshold 
```{r}
ggplot(Bt_BgCor, aes(x=time, y=CorFl, color=sample)) +
  #geom_point(alpha = 1, size = 0.8) +
  geom_smooth()+
  geom_hline(yintercept = AvgTh, linetype = "dashed", color = "grey36", size = 1) +
  annotate("text", x = 60, y=AvgTh, vjust = 1, label = "AvgTh", size = 2.5) +
  geom_hline(yintercept = MaxTh, linetype = "dashed", color = "grey50", size = 1) +
  annotate("text", x= 60 , y=MaxTh, vjust = -1, label = "MaxTh", size = 2.5) +
  ylab("Background corrected Fluorescence (RFU)") +
  xlab("Time (min)")

ggsave(filename = "Bt_Bgcorrected_sensitivity_plot.png", device='png', dpi = 700, width = 16, height = 8)
  #ggtitle("Rs sensitivity Background corrected with 10xSd of NTC-mean")
```

```{r}
ggplot(Sp_test_BgCor, aes(x=time, y=CorFl, color=sample)) +
  #geom_point(alpha = 1, size = 0.8) +
  geom_smooth()+
  geom_hline(yintercept = AvgTh, linetype = "dashed", color = "grey36", size = 1) +
  annotate("text", x = 60, y=AvgTh, vjust = 1, label = "AvgTh", size = 2.5) +
  geom_hline(yintercept = MaxTh, linetype = "dashed", color = "grey50", size = 1) +
  annotate("text", x= 60 , y=MaxTh, vjust = -1, label = "MaxTh", size = 2.5) +
  ylab("Background corrected Fluorescence (RFU)") +
  xlab("Time (min)")

ggsave(filename = "Plamp_Bgcorrected_sensitivity_plot.png", device='png', dpi = 700, width = 16, height = 8)
  #ggtitle("Rs sensitivity Background corrected with 10xSd of NTC-mean")
```

logistic regression of probability of crossing the threshold(0) based on concentration
```{r}
#bull trout first

Bt_lm <- glm(Detect ~ copies.ul, data = gb_Btrt_Ct, family = "binomial")
summary(Bt_lm)
```
plotting the logisitic regression
```{r}
Bt_predict <- data.frame(copies.ul=seq(min(gb_Btrt_Ct$copies.ul), max(gb_Btrt_Ct$copies.ul), len=500))

Bt_predict$detect <- predict(Bt_lm, Bt_predict, type="response")

gb_Bt_logit <- merge(gb_Btrt_Ct, Bt_predict, by="copies.ul")

ggplot(gb_Bt_logit, aes(x=copies.ul, y=detect)) + 
  geom_point() +
  geom_smooth(method = "glm", method.args = list(family="binomial"), se = FALSE)+
  ylab("Probability of detection") +
  xlab("Input concentration (copies/ul)")
```

```{r}
Pl_lm <- glm(Detect ~ copies.ul, data = gb_Plamp_Ct, family = "binomial")
summary(Pl_lm)
```


```{r}
Pl_predict <- data.frame(copies.ul=seq(min(gb_Plamp_Ct$copies.ul), max(gb_Plamp_Ct$copies.ul), len=1000))

Pl_predict$detect <- predict(Pl_lm, Pl_predict, type="response")

gb_Pl_logit <- merge(gb_Plamp_Ct, Pl_predict, by="copies.ul")

ggplot(gb_Pl_logit, aes(x=copies.ul, y=detect)) + 
  geom_point() +
  geom_smooth(method = "glm", method.args = list(family="binomial"), se = FALSE)+
  ylab("Probability of detection") +
  xlab("Input concentration (copies/ul)")
```


